Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-07-04 article-models-ttinternal.tex
preamble cleanup

Let $P$ be a family over an extended context $\ctxext{\Gamma}{A}$. We could
mimic extension by requiring to have terms
\begin{align*}
\jhom*{\Gamma}{{A}{P}}{A}{e_0}\\
\jhom*{{\Gamma}{A}}{{P}{\jcomp{}{e_0}{P}}}{P}{e_1}.
\end{align*}

@@

\begin{defn}
An \emph{extension homomorphism $\cfthom{f}$ from $\extalg{A}$ to
$\extalg{B}$ in context $\Gamma$} is a triple $\unfold{\cfthom{f}}$ consisting of
\begin{align*}
\jhom*
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\cfthomc{\cfthom{f}}}
  \\
\jfhom*
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\cfthomc{\cfthom{f}}}
  {\cftalgf{\cftalg{A}}}
  {\cftalgf{\cftalg{B}}}
  {\cfthomf{\cfthom{f}}}
  \\
\jfhom*
  {\Gamma}
  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
  {\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  {\cftalgt{\cftalg{A}}}
  {\cftalgt{\cftalg{B}}}
  {\cfthomt{\cfthom{f}}}
\end{align*}
for which the diagrams
\begin{equation}\label{eq:exthom1}
\begin{tikzcd}
\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  \ar{d}[swap]{\cftctxext[\cftalg{A}]}
& \ctxext{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}
  \ar{d}{\cftctxext[\cftalg{B}]}
  \\
\cftalgc{\cftalg{A}}
  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
& \cftalgc{\cftalg{B}}
\end{tikzcd}
\end{equation}
and
\begin{equation}\label{eq:exthom2}
\begin{tikzcd}[column sep=huge]
\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
  \ar{r}{\jvcomp{}{\cfthomf{\cfthom{f}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{f}}}}}
  \ar{d}[swap]{\cftfamext[\cftalg{A}]}
& \jcomp{}{\cfthomc{\cfthom{f}}}{\ctxext{\cftalgf{\cftalg{B}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
  \ar{d}{\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamext[\cftalg{B}]}}
  \\
\cftalgf{\cftalg{A}}
  \ar{r}[swap]{\cfthomf{\cfthom{f}}}
& \jcomp{}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}
\end{tikzcd}
\end{equation}
commute judgmentally.
\end{defn}

@@

\begin{lem}\label{lem:empalg-mor}
Consider a family $\jfam{\Gamma}{A}$ with a term $\jterm{\Gamma}{A}{x}$ and 
a family $\jfam{{\Gamma}{A}}{P}$ % with a term $\jterm{{\Gamma}{A}}{P}{y}$
satisfying the judgmental equality
\begin{align}
\jfameq*{\Gamma}{\subst{x}{P}}{A}
  \label{eq1:empalg-mor}
%  \\
%\jtermeq*{\Gamma}{A}{\subst{x}{y}}{x}.
%  \label{eq2:empalg-mor}
\end{align}
Consider also a morphism
\begin{equation*}
\jhom{\Gamma}{{A}{P}}{B}{f}.
\end{equation*}
Then $\subst{\xi}{f}$ is a morphism from $A$ to $B$ 
in context $\Gamma$ and the following inference rules are valid for $\xi$ being 
$x$ or $y$:
\begin{align*}
& \inference
  { \jfam{{\Gamma}{B}}{Q}
    }
  { \jfameq
      {{\Gamma}{A}}
      {\subst{\xi}{\jcomp{}{f}{Q}}}
      {\jcomp{}{\subst{\xi}{f}}{Q}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
      {\subst{\xi}{\jcomp{}{f}{R}}}
      {\jcomp{}{\subst{\xi}{f}}{R}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
    }
  { \jtermeq
      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
      {\jcomp{}{\subst{\xi}{f}}{R}}
      {\subst{\xi}{\jcomp{}{f}{h}}}
      {\jcomp{}{\subst{\xi}{f}}{h}}
    }
\end{align*}
\end{lem}

@@

\begin{defn}
A \emph{CFT-algebra $\cftalg{A}$ in context $\Gamma$} is a septuple
\begin{equation*}
\unfold{\cftalg{A}}
\end{equation*}
where the quintuple $\unfold{\extalg{A}}$ is an extension
algebra in context $\Gamma$, and where
\begin{align*}
\jterm*{\Gamma}{\cftalgc{\cftalg{A}}}{\cftempc{\cftalg{A}}}
  \\
\jterm*{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}
\end{align*}
which satisfy the following judgmental equalities
\begin{enumerate}
\item Families over the empty context in $\cftalg{A}$ are contexts of $\cftalg{A}$:
\begin{align}
\jfameq*
  {\Gamma}
  {\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\cftalgc{\cftalg{A}}}
  \label{empalg-eq1}
\end{align}
\item The empty family over the empty context of $\cftalg{A}$ is the empty context of $\cftalg{A}$:
\begin{align}
\jtermeq*
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\subst{\cftempc{\cftalg{A}}}{\cftempf{\cftalg{A}}}}
  {\cftempc{\cftalg{A}}}.
  \label{empalg-eq2}
\end{align}
\item Context extension of $\cftalg{A}$ is compatible with the empty context and
family of $\cftalg{A}$:
\begin{align}
\jtermeq*
  {{\Gamma}{\cftalgc{\cftalg{A}}}}
  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
  {\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
  {\idtm{\cftalgc{\cftalg{A}}}}
  \label{cftalg-eq1}
  \\
\jtermeq*
  {{\Gamma}{\cftalgc{\cftalg{A}}}}
  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
  {\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
  {\idtm{\cftalgc{\cftalg{A}}}}
  \label{cftalg-eq2}
\end{align}
\item Family extension of $\cftalg{A}$ is compatible with the empty family and the empty family
over families of $\cftalg{A}$:
\begin{align}
\jtermeq*
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
  {\idtm{\cftalgf{\cftalg{A}}}}
  \label{cftalg-eq3}
  \\
\jtermeq*
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
  {\idtm{\cftalgf{\cftalg{A}}}}
  \label{cftalg-eq4}
\end{align}
\item Family extension over the empty context of $\cftalg{A}$ is context extension
of $\cftalg{A}$:
\begin{align}
\jhomeq*
  {\Gamma}
  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\cftalgc{\cftalg{A}}}
  {\subst{\cftempc{\cftalg{A}}}{\cftfamext}}
  {\cftctxext}
  \label{cftalg-eq5}
\end{align}
\end{enumerate}
\end{defn}

@@

\subsubsection{Weakening preserves itself}
Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
  {\cfthom{w}}
\end{equation*}
be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ww} is the
condition that the diagram
\begin{equation}\label{diag:cftwk-cftwk}
\begin{tikzcd}[column sep=10em]
\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}}
  \ar{d}
    [swap]{ \ctxwk
        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
        {\boldsymbol{\mathcal{F}}_{\cfthom{w}}}
      }
  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cfthom{w}}}}
& \ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}
  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}}
  \\
\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cfthomc{\cfthom{w}}}{\cftalg{F_{F_{F_A}}}}}
  \ar{r}[swap]{%
    \jcomp{}
      { \cfthomf{\cfthom{w}}}
      { \jcomp{}
          {\cfthomc{\cfthom{w}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\cfthom{w}}
                }
            }
        }
    }
& \jcomp{}
    {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
    {\jcomp{}{\cfthomc{\cfthom{w}}}{\cftalg{F_{F_{F_A}}}}}
\end{tikzcd}
\end{equation}
of CFT-algebras in context 
$ \ctxext
    { { { {\Gamma}
          {\cftalgc{\cftalg{A}}}
          }
        { \cftalgf{\cftalg{A}}}
        }
      { \ctxwk
          {\cftalgf{\cftalg{A}}}
          {\cftalgf{\cftalg{A}}}
        }
      }
    { \ctxwk
        {\cftalgf{\cftalg{A}}}
        {\cftalgf{\cftalg{F_A}}}
      }$ %
commutes. In this diagram, it is not immediately obvious that 
\begin{enumerate}
\item the domain of the CFT-homomorphism
$\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}$
is the indicated domain. Note that we have the judgmental equalities
\begin{align*}
\jcomp{}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
& \jdeq
\ctxwk
  {\cftalgf{\cftalg{A}}}
  {\jcomp{}{\cftfamext}{\cftalg{F_{F_A}}}}
  \\
& \jdeq
\ctxwk
  {\cftalgf{\cftalg{A}}}
  {\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftalg{F_{F_A}}}}
  \\
& \jdeq
\ctxwk
  {\cftalgf{\cftalg{A}}}
  {\cftalg{F_{F_{F_A}}}}.
\end{align*}
\item the domain and codomain of the CFT-homomorphism
\begin{equation*}
\jcomp{}
      { \cfthomf{\cfthom{w}}}
      { \jcomp{}
          {\cfthomc{\cfthom{w}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\cfthom{w}}
                }
            }
        }
\end{equation*}
from
$ \jcomp{}
      { \cfthomf{\cfthom{w}}}
      { \jcomp{}
          {\cfthomc{\cfthom{w}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
                }
            }
        }
$ to
$ \jcomp{}
      { \cfthomf{\cfthom{w}}}
      { \jcomp{}
          {\cfthomc{\cfthom{w}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\cftalg{F_{F_A}}}
                }
            }
        }
$ are indeed judgmentally equal to the indicated domain and codomain. To see
this, we have the following lemma:
\end{enumerate}

@@

\begin{lem}
Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
  {\cfthom{w}}
\end{equation*}
be a homomorphism of CFT-algebras for which the diagram in
\autoref{diag:cftwk-cftwk} commutes judgmentally. Then
we have the judgmental equality
\begin{align*}
& \jcomp{}
  { \subst{q}{{\apply\cftctxext{\gamma}{b}}{\cfthom{w}}}
    }
  { \subst
      {\apply\cftfamext{\gamma}{b}{q}}
      {{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}}}
    }
  \\
& \jdeq
  \jcomp{}
    { \subst{b}{{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cfthom{w}}}}}
      }
    { \subst
        { {q}{{b}{{a}{{\gamma}{\cfthomf{\cfthom{w}}}}}}}
        { {\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{\subst{b}{{a}{{\gamma}{\cfthomc{\cfthom{w}}}}}}}
          {\cfthom{w}}
          }
      }
\end{align*}
of CFT-homomorphisms from $\subst{b}{{\gamma}{\cftalg{F_{F_A}}}}$ to
$ \subst
      { {q}{{b}{{a}{{\gamma}{\cfthomf{\cfthom{w}}}}}}
        }
      { {{b}{{a}{{\gamma}{\cfthomc{\cfthom{w}}}}}}
        {{a}{{\gamma}{\cftalg{F_{F_{F_{F_{A}}}}}}}}
        }$
in context $\Gamma$, for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
$q:\subst{b}{{\gamma}{\cftalgf{\cftalg{F_A}}}}$.
\end{lem}

>>>>

Let $P$ be a family over an extended context $\ctxext{\Gamma}{A}$. We could
mimic extension by requiring to have terms
\begin{align*}
\jalign\jhom{\Gamma}{{A}{P}}{A}{e_0}\\
\jalign\jhom{{\Gamma}{A}}{{P}{\jcomp{}{e_0}{P}}}{P}{e_1}.
\end{align*}

@@

\begin{defn}
An \emph{extension homomorphism $\cfthom{f}$ from $\extalg{A}$ to
$\extalg{B}$ in context $\Gamma$} is a triple $\unfold{\cfthom{f}}$ consisting of
\begin{align*}
\jalign\jhom
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\cfthomc{\cfthom{f}}}
  \\
\jfhom*
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\cfthomc{\cfthom{f}}}
  {\cftalgf{\cftalg{A}}}
  {\cftalgf{\cftalg{B}}}
  {\cfthomf{\cfthom{f}}}
  \\
\jfhom*
  {\Gamma}
  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
  {\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  {\cftalgt{\cftalg{A}}}
  {\cftalgt{\cftalg{B}}}
  {\cfthomt{\cfthom{f}}}
\end{align*}
for which the diagrams
\begin{equation}\label{eq:exthom1}
\begin{tikzcd}
\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  \ar{d}[swap]{\cftctxext[\cftalg{A}]}
& \ctxext{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}
  \ar{d}{\cftctxext[\cftalg{B}]}
  \\
\cftalgc{\cftalg{A}}
  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
& \cftalgc{\cftalg{B}}
\end{tikzcd}
\end{equation}
and
\begin{equation}\label{eq:exthom2}
\begin{tikzcd}[column sep=huge]
\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
  \ar{r}{\jvcomp{}{\cfthomf{\cfthom{f}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{f}}}}}
  \ar{d}[swap]{\cftfamext[\cftalg{A}]}
& \jcomp{}{\cfthomc{\cfthom{f}}}{\ctxext{\cftalgf{\cftalg{B}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
  \ar{d}{\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamext[\cftalg{B}]}}
  \\
\cftalgf{\cftalg{A}}
  \ar{r}[swap]{\cfthomf{\cfthom{f}}}
& \jcomp{}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}
\end{tikzcd}
\end{equation}
commute judgmentally.
\end{defn}

@@

\begin{lem}\label{lem:empalg-mor}
Consider a family $\jfam{\Gamma}{A}$ with a term $\jterm{\Gamma}{A}{x}$ and 
a family $\jfam{{\Gamma}{A}}{P}$ % with a term $\jterm{{\Gamma}{A}}{P}{y}$
satisfying the judgmental equality
\begin{align}
\jalign\jfameq{\Gamma}{\subst{x}{P}}{A}
  \label{eq1:empalg-mor}
%  \\
%\jalign\jtermeq{\Gamma}{A}{\subst{x}{y}}{x}.
%  \label{eq2:empalg-mor}
\end{align}
Consider also a morphism
\begin{equation*}
\jhom{\Gamma}{{A}{P}}{B}{f}.
\end{equation*}
Then $\subst{\xi}{f}$ is a morphism from $A$ to $B$ 
in context $\Gamma$ and the following inference rules are valid for $\xi$ being 
$x$ or $y$:
\begin{align*}
& \inference
  { \jfam{{\Gamma}{B}}{Q}
    }
  { \jfameq
      {{\Gamma}{A}}
      {\subst{\xi}{\jcomp{}{f}{Q}}}
      {\jcomp{}{\subst{\xi}{f}}{Q}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
      {\subst{\xi}{\jcomp{}{f}{R}}}
      {\jcomp{}{\subst{\xi}{f}}{R}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{B}}{Q}}{R}{h}
    }
  { \jtermeq
      {{{\Gamma}{A}}{\jcomp{}{\subst{\xi}{f}}{Q}}}
      {\jcomp{}{\subst{\xi}{f}}{R}}
      {\subst{\xi}{\jcomp{}{f}{h}}}
      {\jcomp{}{\subst{\xi}{f}}{h}}
    }
\end{align*}
\end{lem}

@@

\begin{defn}
A \emph{CFT-algebra $\cftalg{A}$ in context $\Gamma$} is a septuple
\begin{equation*}
\unfold{\cftalg{A}}
\end{equation*}
where the quintuple $\unfold{\extalg{A}}$ is an extension
algebra in context $\Gamma$, and where
\begin{align*}
\jalign\jterm{\Gamma}{\cftalgc{\cftalg{A}}}{\cftempc{\cftalg{A}}}
  \\
\jalign\jterm{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}
\end{align*}
which satisfy the following judgmental equalities
\begin{enumerate}
\item Families over the empty context in $\cftalg{A}$ are contexts of $\cftalg{A}$:
\begin{align}
\jalign\jfameq
  {\Gamma}
  {\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\cftalgc{\cftalg{A}}}
  \label{empalg-eq1}
\end{align}
\item The empty family over the empty context of $\cftalg{A}$ is the empty context of $\cftalg{A}$:
\begin{align}
\jtermeq*
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\subst{\cftempc{\cftalg{A}}}{\cftempf{\cftalg{A}}}}
  {\cftempc{\cftalg{A}}}.
  \label{empalg-eq2}
\end{align}
\item Context extension of $\cftalg{A}$ is compatible with the empty context and
family of $\cftalg{A}$:
\begin{align}
\jtermeq*
  {{\Gamma}{\cftalgc{\cftalg{A}}}}
  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
  {\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
  {\idtm{\cftalgc{\cftalg{A}}}}
  \label{cftalg-eq1}
  \\
\jtermeq*
  {{\Gamma}{\cftalgc{\cftalg{A}}}}
  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
  {\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
  {\idtm{\cftalgc{\cftalg{A}}}}
  \label{cftalg-eq2}
\end{align}
\item Family extension of $\cftalg{A}$ is compatible with the empty family and the empty family
over families of $\cftalg{A}$:
\begin{align}
\jtermeq*
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
  {\idtm{\cftalgf{\cftalg{A}}}}
  \label{cftalg-eq3}
  \\
\jtermeq*
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
  {\idtm{\cftalgf{\cftalg{A}}}}
  \label{cftalg-eq4}
\end{align}
\item Family extension over the empty context of $\cftalg{A}$ is context extension
of $\cftalg{A}$:
\begin{align}
\jhomeq*
  {\Gamma}
  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\cftalgc{\cftalg{A}}}
  {\subst{\cftempc{\cftalg{A}}}{\cftfamext}}
  {\cftctxext}
  \label{cftalg-eq5}
\end{align}
\end{enumerate}
\end{defn}

@@

\subsubsection{Weakening preserves itself}
Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
  {\cftwk{\cftalg{A}}}
\end{equation*}
be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ww} is the
condition that the diagram
\begin{equation}\label{diag:cftwk-cftwk}
\begin{tikzcd}[column sep=10em]
\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}}
  \ar{d}
    [swap]{ \ctxwk
        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
        {\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}
      }
  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
& \ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}
  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}
  \\
\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}{\jcomp{}{\cftwkc{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}}
  \ar{r}[swap]{%
    \jcomp{}
      { \cftwkf{\cftalg{A}}}
      { \jcomp{}
          {\cftwkc{\cftalg{A}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\cftwk{\cftalg{A}}}
                }
            }
        }
    }
& \jcomp{}
    {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
    {\jcomp{}{\cftwkc{\cftalg{A}}}{\cftalg{F_{F_{F_A}}}}}
\end{tikzcd}
\end{equation}
of CFT-algebras in context 
$ \ctxext
    { { { {\Gamma}
          {\cftalgc{\cftalg{A}}}
          }
        { \cftalgf{\cftalg{A}}}
        }
      { \ctxwk
          {\cftalgf{\cftalg{A}}}
          {\cftalgf{\cftalg{A}}}
        }
      }
    { \ctxwk
        {\cftalgf{\cftalg{A}}}
        {\cftalgf{\cftalg{F_A}}}
      }$ %
commutes. In this diagram, it is not immediately obvious that 
\begin{enumerate}
\item the domain of the CFT-homomorphism
$\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}$
is the indicated domain. Note that we have the judgmental equalities
\begin{align*}
\jcomp{}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
& \jdeq
\ctxwk
  {\cftalgf{\cftalg{A}}}
  {\jcomp{}{\cftfamext}{\cftalg{F_{F_A}}}}
  \\
& \jdeq
\ctxwk
  {\cftalgf{\cftalg{A}}}
  {\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftalg{F_{F_A}}}}
  \\
& \jdeq
\ctxwk
  {\cftalgf{\cftalg{A}}}
  {\cftalg{F_{F_{F_A}}}}.
\end{align*}
\item the domain and codomain of the CFT-homomorphism
\begin{equation*}
\jcomp{}
      { \cftwkf{\cftalg{A}}}
      { \jcomp{}
          {\cftwkc{\cftalg{A}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\cftwk{\cftalg{A}}}
                }
            }
        }
\end{equation*}
from
$ \jcomp{}
      { \cftwkf{\cftalg{A}}}
      { \jcomp{}
          {\cftwkc{\cftalg{A}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
                }
            }
        }
$ to
$ \jcomp{}
      { \cftwkf{\cftalg{A}}}
      { \jcomp{}
          {\cftwkc{\cftalg{A}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\cftalg{F_{F_A}}}
                }
            }
        }
$ are indeed judgmentally equal to the indicated domain and codomain. To see
this, we have the following lemma:
\end{enumerate}

@@

\begin{lem}
Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
  {\cftwk{\cftalg{A}}}
\end{equation*}
be a homomorphism of CFT-algebras for which the diagram in
\autoref{diag:cftwk-cftwk} commutes judgmentally. Then
we have the judgmental equality
\begin{align*}
& \jcomp{}
  { \subst{q}{{\apply\cftctxext{\gamma}{b}}{\cftwk{\cftalg{A}}}}
    }
  { \subst
      {\apply\cftfamext{\gamma}{b}{q}}
      {{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
    }
  \\
& \jdeq
  \jcomp{}
    { \subst{b}{{a}{{\gamma}{\boldsymbol{\mathcal{F}}_{\cftwk{\cftalg{A}}}}}}
      }
    { \subst
        { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}}
        { {\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{\subst{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}}
          {\cftwk{\cftalg{A}}}
          }
      }
\end{align*}
of CFT-homomorphisms from $\subst{b}{{\gamma}{\cftalg{F_{F_A}}}}$ to
$ \subst
      { {q}{{b}{{a}{{\gamma}{\cftwkf{\cftalg{A}}}}}}
        }
      { {{b}{{a}{{\gamma}{\cftwkc{\cftalg{A}}}}}}
        {{a}{{\gamma}{\cftalg{F_{F_{F_{F_{A}}}}}}}}
        }$
in context $\Gamma$, for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
$q:\subst{b}{{\gamma}{\cftalgf{\cftalg{F_A}}}}$.
\end{lem}

\subsubsection{Currying for weakening}
Let $\cftalg{A}$ be a CFT-algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalg{F_A}}}
  {\jcomp{}{\cftctxext}{\cftalg{F_A}}}
  {\cftwk{\cftalg{A}}}
\end{equation*}
be a CFT-homomorphism. The abstraction of the rules in \autoref{comp-ew} is the
condition that the diagram
\begin{equation}\label{eq:cftwk-cftextcurry}
\begin{tikzcd}[column sep=large]
\ctxwk
  {\ctxext{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{F_A}}}}
  {\cftalg{F_A}}
  \ar{r}{\ctxwk{\cftalgf{\cftalg{F_A}}}{\cftwk{\cftalg{A}}}}
  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
& \ctxwk{\cftalgf{\cftalg{F_A}}}{\cftalg{F_{F_A}}}
  \ar{d}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \\
& \cftalg{F_{F_{F_A}}}
\end{tikzcd}
\end{equation}
of CFT-algebras in context 
$\ctxext
    {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
    {\cftalgf{\cftalg{F_A}}}
$ %
commutes.
